GBM Recurrence - Mine Single Cell data

### 2023-10-24 reports/gbm_recurrence_mine_single_cell_data_preprocessing.qmd
revision 2 commit

@@

# matrix data

matrix_ncbi <- read_csv(here::here("data/revision_single_cell_data/from_ncbi/GSM4972210_Human.GBM.R1_2_3_4_4nc.filtered.gene.bc.matrix.csv"))
```

```{r}
View(annot_1)
#View(annot_2_ncbi)
#View(matrix_ncbi)
```

@@

@@

```{r}
cpm_mat <- calculateCPM(matrix_ncbi)
mat_cpm_log2 <- log2(cpm_mat + 1)
rm(cpm_mat)
```

```{r}
mat_cpm_log2[1:15, 1:15]
```

@@

```{r}
mat_cpm_df <- mat_cpm_log2 %>% 
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  relocate(gene, everything()) 
```

@@

```{r}
#View(long_log2_df_asah1_annot)
```

@@

>>>>

annot_2_nd <- read_csv(here::here("data/revision_single_cell_data/from_ncbi/GSM4972211_annot.Human.GBM.ND1_2_3_4_5_6_7.csv"))

@@

# matrix data
matrix_ncbi <- read_csv(here::here("data/revision_single_cell_data/from_ncbi/GSM4972210_Human.GBM.R1_2_3_4_4nc.filtered.gene.bc.matrix.csv"))

# matrix data init
matrix_ncbi_nd <- read_csv(here::here("data/revision_single_cell_data/from_ncbi/GSM4972211_Human.GBM.ND1_2_3_4_5_6_7.filtered.gene.bc.matrix.csv"))
```

```{r}
#View(annot_1)
#View(annot_2_ncbi)
#View(matrix_ncbi)
```

@@

## rGBM

@@

```{r}
if(!file.exists(here::here("rds/matrix_ncbi_scSeq_CPM_rgbm.rds"))) {

  cpm_mat <- calculateCPM(matrix_ncbi)
  mat_cpm_log2 <- log2(cpm_mat + 1)
  rm(cpm_mat)

  saveRDS(mat_cpm_log2, 
          here::here("rds/matrix_ncbi_scSeq_CPM_rgbm.rds"))
} else {

  mat_cpm_log2 <- readRDS(here::here("rds/matrix_ncbi_scSeq_CPM_rgbm.rds"))

}
```

```{r}
mat_cpm_log2[10:15, 10:15]
```

## iGBM (ND)

```{r}
matrix_ncbi_nd <- matrix_ncbi_nd %>%
  dplyr::rename('gene' = '...1') %>%
  column_to_rownames("gene") %>% 
  as.matrix()
```

```{r}
if(!file.exists(here::here("rds/matrix_ncbi_scSeq_CPM_igbm.rds"))) {

  cpm_mat_nd <- calculateCPM(matrix_ncbi_nd)
  mat_cpm_log2_nd <- log2(cpm_mat_nd + 1)
  rm(cpm_mat_nd)

  saveRDS(mat_cpm_log2_nd, 
          here::here("rds/matrix_ncbi_scSeq_CPM_igbm.rds"))
} else {

  mat_cpm_log2 <- readRDS(here::here("rds/matrix_ncbi_scSeq_CPM_igbm.rds"))

}
```

```{r}
mat_cpm_log2_nd[10:15, 10:15]
```

@@

## rGBM

```{r}
mat_cpm_df <- mat_cpm_log2 %>% 
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  relocate(gene, everything())

mat_cpm_df_names <- mat_cpm_df %>%
  dplyr::select(gene)
```

## iGBM (ND)

```{r}
mat_cpm_df_nd <- mat_cpm_log2_nd %>% 
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  relocate(gene, everything())

mat_cpm_df_names_nd <- mat_cpm_df_nd %>%
  dplyr::select(gene)

  #View(mat_cpm_df_nd)
```

# Filter for ASAH1

## rGBM

@@

```{r}
View(long_log2_df_asah1_annot)
```

@@

## iGBM (ND)

```{r}
# ASAH1 expression
mat_log2_df_asah1_nd <- mat_cpm_df_nd %>%
  filter(gene == "ASAH1")
```

```{r}
long_log2_df_asah1_nd <- mat_log2_df_asah1_nd %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "asah1_expression") 
```

```{r}
long_log2_df_asah1_annot_nd <- long_log2_df_asah1_nd %>%
  left_join(annot_2_nd, by = c("cell" = "cell"))
```

```{r}
saveRDS(long_log2_df_asah1_annot_nd,
        here::here("rds/long_cpm_scrnaseq_asah1_annot_nd.rds"))
```

# Filter for MPO

## rGBM

```{r}
# MPO expression
mat_log2_df_mpo <- mat_cpm_df %>%
  filter(gene == "MPO")
```

```{r}
long_log2_df_mpo <- mat_log2_df_mpo %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "mpo_expression") 
```

```{r}
long_log2_df_mpo_annot <- long_log2_df_mpo %>%
  left_join(annot_1, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_mpo_annot)
```

```{r}
saveRDS(long_log2_df_mpo_annot,
        here::here("rds/long_cpm_scrnaseq_mpo_annot.rds"))
```

## iGBM (ND)

```{r}
# MPO expression

mat_log2_df_mpo_nd <- mat_cpm_df_nd %>%
  filter(gene == "MPO")
```

```{r}
long_log2_df_mpo_nd <- mat_log2_df_mpo_nd %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "mpo_expression") 
```

```{r}
long_log2_df_mpo_annot_nd <- long_log2_df_mpo_nd %>%
  left_join(annot_2_nd, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_mpo_annot_nd)
```

```{r}
saveRDS(long_log2_df_mpo_annot_nd,
        here::here("rds/long_cpm_scrnaseq_mpo_annot_nd.rds"))
```

# Filter for CD14

## rGBM

```{r}
# CD14 expression
mat_log2_df_cd14 <- mat_cpm_df %>%
  filter(gene == "CD14")
```

```{r}
long_log2_df_cd14 <- mat_log2_df_cd14 %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "cd14_expression") 
```

```{r}
long_log2_df_cd14_annot <- long_log2_df_cd14 %>%
  left_join(annot_1, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_cd14_annot)
```

```{r}
saveRDS(long_log2_df_cd14_annot,
        here::here("rds/long_cpm_scrnaseq_cd14_annot.rds"))
```

## iGBM (ND)

```{r}
# CD14 expression
mat_log2_df_cd14_nd <- mat_cpm_df_nd %>%
  filter(gene == "CD14")
```

```{r}
long_log2_df_cd14_nd <- mat_log2_df_cd14_nd %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "cd14_expression") 
```

```{r}
long_log2_df_cd14_annot_nd <- long_log2_df_cd14_nd %>%
  left_join(annot_2_nd, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_cd14_annot_nd)
```

```{r}
saveRDS(long_log2_df_cd14_annot_nd,
        here::here("rds/long_cpm_scrnaseq_cd14_annot_nd.rds"))
```

# Filter for IBA1

## rGBM

```{r}
# IBA1 expression
mat_log2_df_iba1 <- mat_cpm_df %>%
  filter(gene == "AIF1")
```

```{r}
long_log2_df_iba1 <- mat_log2_df_iba1 %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "iba1_expression") 
```

```{r}
long_log2_df_iba1_annot <- long_log2_df_iba1 %>%
  left_join(annot_1, by = c("cell" = "cell"))
```

```{r}
saveRDS(long_log2_df_iba1_annot,
        here::here("rds/long_cpm_scrnaseq_iba1_annot.rds"))
```

## iGBM (ND)

```{r}
# IBA1 expression
mat_log2_df_iba1_nd <- mat_cpm_df_nd %>%
  filter(gene == "AIF1")
```

```{r}
long_log2_df_iba1_nd <- mat_log2_df_iba1_nd %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "iba1_expression") 
```

```{r}
long_log2_df_iba1_annot_nd <- long_log2_df_iba1_nd %>%
  left_join(annot_2_nd, by = c("cell" = "cell"))
```

```{r}
saveRDS(long_log2_df_iba1_annot_nd,
        here::here("rds/long_cpm_scrnaseq_iba1_annot_nd.rds"))
```

# Filter for lipocalin-2 (LCN2)

## rGBM

```{r}
mat_cpm_df[1:5, 1:5]
```

```{r}
# LCN2 expression
mat_log2_df_lcn2 <- mat_cpm_df %>%
  filter(gene == "LCN2")
```

```{r}
long_log2_df_lcn2 <- mat_log2_df_lcn2 %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "lcn2_expression") 
```

```{r}
long_log2_df_lcn2_annot <- long_log2_df_lcn2 %>%
  left_join(annot_1, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_lcn2_annot)
head(long_log2_df_lcn2_annot)
```

```{r}
saveRDS(long_log2_df_lcn2_annot,
        here::here("rds/long_cpm_scrnaseq_lcn2_annot.rds"))
```

## iGBM (ND)

```{r}
# LCN2 expression
mat_log2_df_lcn2_nd <- mat_cpm_df_nd %>%
  filter(gene == "LCN2")
```

```{r}
long_log2_df_lcn2_nd <- mat_log2_df_lcn2_nd %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "lcn2_expression") 
```

```{r}
long_log2_df_lcn2_annot_nd <- long_log2_df_lcn2_nd %>%
  left_join(annot_2_nd, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_lcn2_annot_nd)
head(long_log2_df_lcn2_annot_nd)
```

```{r}
saveRDS(long_log2_df_lcn2_annot_nd,
        here::here("rds/long_cpm_scrnaseq_lcn2_annot_nd.rds"))
```

# Filter for neutrophil elastase (ELANE)

## rGBM

```{r}
# ELANE expression
mat_log2_df_elane <- mat_cpm_df %>%
  filter(gene == "ELANE")
```

```{r}
long_log2_df_elane <- mat_log2_df_elane %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "elane_expression") 
```

```{r}
long_log2_df_elane_annot <- long_log2_df_elane %>%
  left_join(annot_1, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_elane_annot)
head(long_log2_df_elane_annot)
```

```{r}
saveRDS(long_log2_df_elane_annot,
        here::here("rds/long_cpm_scrnaseq_elane_annot.rds"))
```

## iGBM (ND)

```{r}
# ELANE expression
mat_log2_df_elane_nd <- mat_cpm_df_nd %>%
  filter(gene == "ELANE")
```

```{r}
long_log2_df_elane_nd <- mat_log2_df_elane_nd %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "elane_expression") 
```

```{r}
long_log2_df_elane_annot_nd <- long_log2_df_elane_nd %>%
  left_join(annot_2_nd, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_elane_annot_nd)
head(long_log2_df_elane_annot_nd)
```

```{r}
saveRDS(long_log2_df_elane_annot_nd,
        here::here("rds/long_cpm_scrnaseq_elane_annot_nd.rds"))
```

# Filter for CD68

## rGBM

```{r}
# CD68 expression
mat_log2_df_cd68 <- mat_cpm_df %>%
  filter(gene == "CD68")
```

```{r}
long_log2_df_cd68 <- mat_log2_df_cd68 %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "cd68_expression") 
```
Merge with annotation

```{r}
long_log2_df_cd68_annot <- long_log2_df_cd68 %>%
  left_join(annot_1, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_cd68_annot)
head(long_log2_df_cd68_annot)
```

```{r}
saveRDS(long_log2_df_cd68_annot,
        here::here("rds/long_cpm_scrnaseq_cd68_annot.rds"))
```

## iGBM (ND)

```{r}
# CD68 expression
mat_log2_df_cd68_nd <- mat_cpm_df_nd %>%
  filter(gene == "CD68")
```

```{r}
long_log2_df_cd68_nd <- mat_log2_df_cd68_nd %>%
  pivot_longer(cols = -gene, 
               names_to = "cell", 
               values_to = "cd68_expression") 
```

```{r}
long_log2_df_cd68_annot_nd <- long_log2_df_cd68_nd %>%
  left_join(annot_2_nd, by = c("cell" = "cell"))
```

```{r}
#View(long_log2_df_cd68_annot_nd)
head(long_log2_df_cd68_annot_nd)
```

```{r}
saveRDS(long_log2_df_cd68_annot_nd,
        here::here("rds/long_cpm_scrnaseq_cd68_annot_nd.rds"))
```
